Analysis: Exploratory Data Analysis

### 2021-11-11 code/analysis_code/analysis-eda.Rmd
EDA complete, linear analysis started

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
* summarytools: for overall dataframe summary
* ggplot2: for plotting data
* car: for creating QQ plots
* gt: for creating tables for summary statistics / other numerical outputs
* scales: for percent calculation

@@

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(summarytools) #for overall dataframe summary
library(ggplot2) #for plotting data
library(car) #for creating QQ plots
library(gt) #for creating tables
library(scales) #for percent calculation

@@

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
summarytools::dfSummary(EDAdata)
```

@@

## Outcome of Interest: Requested FEMA Funds
The first of two primary outcomes of interest is the total amount of funds the state requested from FEMA for a disaster declaration.
```{r}
#summary statistics
base::summary(EDAdata$ReqAmt)

@@

#boxplot
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot()

#crop to $1M
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot() +
  ylim(0, 1000000)

##QQ plot
car::qqPlot(EDAdata$ReqAmt)
```

## Outcome of Interest: Obligated FEMA Funds

>>>>

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
* summarytools: for overall dataframe summary
* ggplot2: for plotting data
* car: for creating QQ plots
* gtsummary: for creating tables for summary statistics / other numerical outputs
* scales: for percent calculation and axis formatting
* cowplot: for placing plots side by side
* table1: for making summary tables

@@

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(summarytools) #for overall dataframe summary
library(ggplot2) #for plotting data
library(car) #for creating QQ plots
library(gtsummary) #for creating tables
library(scales) #for percent calculation
library(cowplot) #for placing plots side by side
library(table1) #for making summary tables

@@

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
summarytools::dfSummary(EDAdata)
```
Looking at the output of `summarytools::dfsummary()`

* There are 445 distinct disaster numbers across 2012 - 2021. 
* `OblAmt` ranges from -\$855,200.3 to \$4,408,424,297.
* `ReqAmt` ranges from -\$855,200.3 to \$4,388,540,486.
* `ResponseDays` are missing 72% of the time, which suggests it should not be included in the analysis.
* There are 16 possible incident types
* There is a clear effect of both incident month and incident year.

Let's dive a little deeper into each variable.

@@

## Outcome of Interest: Requested FEMA Funds
The first of two primary outcomes of interest is the total amount of funds the state requested from the federal government for a disaster declaration.
```{r}
#summary statistics
base::summary(EDAdata$ReqAmt)

@@

#try a log transformation out of curiosity 
#this does ignore all negative values
ggplot2::ggplot(data = EDAdata, aes(x = log(ReqAmt))) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot()

#crop to $1M to get a better idea
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot() +
  ylim(0, 1000000)

#log transformed boxplot
ggplot2::ggplot(data = EDAdata, aes(y = log(ReqAmt))) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$ReqAmt)
```
These outputs show the requested amounts do not follow a normal distribution, which is to be expected given the broad range of disasters. We may need to remove outliers in the analysis, but for now, we'll leave them to capture the true nature of disaster declarations. The largest outlier is likely the COVID-19 Pandemic. The log transformation does a better job of showing the distribution of the data, so we may need to consider an exponential model in the analysis; however, log(negative number) is undefined, so we would be ignoring all reimbursement data. The transformed and untransformed figures suggest there are significant tails in the data.

## Outcome of Interest: Obligated FEMA Funds
The alternative outcome of interest is the funding obligated from FEMA. While this may be the same as `ReqAmt` (especially for large-scale disasters), FEMA doesn't always pay the amount the states request.
```{r obligated funds}
#summary statistics
base::summary(EDAdata$OblAmt)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = OblAmt)) +
  geom_density()

#try a log transformation out of curiosity 
ggplot2::ggplot(data = EDAdata, aes(x = log(OblAmt))) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = OblAmt)) +
  geom_boxplot()

#crop to $1M to get a better idea
ggplot2::ggplot(data = EDAdata, aes(y = OblAmt)) +
  geom_boxplot() +
  ylim(0, 1000000)

#log transformed boxplot
ggplot2::ggplot(data = EDAdata, aes(y = log(OblAmt))) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$OblAmt)
```
It appears that obligated funds and requested funds have similar distributions and characteristics. The same commentary about the requested funds analysis applies here as well.

## Predictor: State
```{r}
#frequency table since it's categorical
#hide NAs because there are no NAs
summarytools::freq(EDAdata$state, report.nas = FALSE)

#examine graphical relationship with outcomes
#include a jitter function to have a better idea of number of measurements and distribution
#filter top 5 states
#start with obligated amount
obl_fig_state <- EDAdata %>%
                    dplyr::add_count(state) %>%
                    dplyr::filter(dplyr::dense_rank(-n) < 6) %>%
                    ggplot2::ggplot(aes(x = state, y = OblAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "tomato") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      labs(x = "State or Territory",
                           y = "Obligated FEMA Funds",
                           title = "FEMA funds obligated to five most disaster-prone states")
obl_fig_state

#save file
obl_state_fig_file = here("results","obl-state.png")
ggsave(filename = obl_state_fig_file, plot = obl_fig_state)

#repeat for requested funds
req_fig_state <- EDAdata %>%
                    dplyr::add_count(state) %>%
                    dplyr::filter(dplyr::dense_rank(-n) < 6) %>%
                    ggplot2::ggplot(aes(x = state, y = ReqAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "turquoise4") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      labs(x = "State or Territory",
                           y = "Requested FEMA Funds",
                           title = "FEMA funds requested by five most disaster-prone states")
req_fig_state

#save file
req_state_fig_file = here("results","req-state.png")
ggsave(filename = req_state_fig_file, plot = req_fig_state)

#create a table where columns represent requested and obligated funds and rows are 5 most disaster-prone states
decs_state_5 <- EDAdata %>%
                    dplyr::add_count(state) %>%
                    dplyr::filter(dense_rank(-n) < 6) %>%
                    gtsummary::select(state, ReqAmt, OblAmt) %>%
                    gtsummary::tbl_summary(
                      by = state,
                      label = list(state ~ "State or Territory",
                                   ReqAmt ~ "Requested Funds",
                                   OblAmt ~ "Obligated Funds"),
                      statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                    gtsummary::modify_header(label ~ "**Variable**") %>% 
                    gtsummary::modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5") ~ "**State**") %>%
                    gtsummary::modify_caption("**Table 1. FEMA Funding for 5 Most Disaster-Prone States 2012 - 2021**") %>%
                    gtsummary::bold_labels()
decs_state_5

#save file
decs_state_file = here("results", "decs-state-5.Rds")
saveRDS(decs_state_5, file = decs_state_file)
```

## Predictor: Incident Year
```{r}
#summary statistics since it's continuous
base::summary(EDAdata$IncidentYear)

#examine graphical relationship with outcomes

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = IncidentYear)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = IncidentYear)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$IncidentYear)

#histogram distribution
year_hist <- EDAdata %>% 
               ggplot2::ggplot(aes(x=IncidentYear)) + 
                        geom_bar(fill="steelblue4") + 
                        scale_x_continuous(breaks = seq(2012, 2021, by = 1)) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 125)) +
                        geom_text(stat='count', aes(label=..count..), vjust=-1) +
                        labs(x = "Incident Year",
                             y = "Number of Declarations",
                             title = "Disaster Declarations per Year (2012 - 2021)")
year_hist

#save file
year_fig_file = here("results","incidentyear.png")
ggsave(filename = year_fig_file, plot = year_hist)

#examine graphical relationship with outcomes of interest
#look at total cost per incident year
req_fig_year <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentYear)) +
                      geom_bar(aes(y = ReqAmt),
                               stat = "identity",
                               fill = "tomato") +
                      scale_y_continuous(expand = c(0, 0), 
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(breaks = seq(2012, 2021, by = 1)) +
                      labs(x = "Incident Year",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_year

#repeat for obligated funds
obl_fig_year <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentYear)) +
                      geom_bar(aes(y = OblAmt),
                               stat = "identity",
                               fill = "turquoise4") +
                      scale_y_continuous(expand = c(0, 0),
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(breaks = seq(2012, 2021, by = 1)) +
                      labs(x = "Incident Year",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_year

#combine into one plot
plot_col <- cowplot::plot_grid(req_fig_year, 
                                     obl_fig_year,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col

title1 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds per year",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_year <- cowplot::plot_grid(
                              title1, plot_col,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_year

#save file
funds_year_fig_file = here("results","funds-years.png")
ggsave(filename = funds_year_fig_file, plot = funds_fig_year)

#create a table where columns represent requested and obligated funds and rows are 5 most disaster-prone states
decs_year_3 <- EDAdata %>%
                    dplyr::add_count(IncidentYear) %>%
                    dplyr::filter(dense_rank(-n) < 4) %>%
                    gtsummary::select(IncidentYear, ReqAmt, OblAmt) %>%
                    gtsummary::tbl_summary(
                      by = IncidentYear,
                      label = list(IncidentYear ~ "Year",
                                   ReqAmt ~ "Requested Funds",
                                   OblAmt ~ "Obligated Funds"),
                      statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                    gtsummary::modify_header(label ~ "**Variable**") %>% 
                    gtsummary::modify_spanning_header(c("stat_1", "stat_2", "stat_3") ~ "**Year**") %>%
                    gtsummary::modify_caption("**Table 2. FEMA Funding for the 3 Costliest Years (2012 - 2021)**") %>%
                    gtsummary::bold_labels()
decs_year_3

#save file
decs_year_file = here("results", "decs-year-3.Rds")
saveRDS(decs_year_3, file = decs_year_file)
```

## Incident Month
```{r}
#summary statistics since it's continuous
base::summary(EDAdata$IncidentMonth)

#examine graphical relationship with outcomes

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = IncidentMonth)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = IncidentMonth)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$IncidentMonth)

#histogram distribution
month_hist <- EDAdata %>% 
               ggplot2::ggplot(aes(x=IncidentMonth)) + 
                        geom_bar(fill="darkseagreen4") + 
                        scale_x_continuous(breaks = seq(1, 12, by = 1),
                                           labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                                      "Jul", "Aug", "Sept", "Oct", "Nov", "Dec")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 110)) +
                        geom_text(stat='count', aes(label=..count..), vjust=-1) +
                        labs(x = "Declaration Month",
                             y = "Number of Declarations",
                             title = "Disaster Declarations per Month (2012 - 2021)")
month_hist

#save file
month_fig_file = here("results","incidentmonth.png")
ggsave(filename = month_fig_file, plot = month_hist)

#examine graphical relationship with outcomes of interest
#look at total cost per incident month
req_fig_month <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentMonth)) +
                      geom_bar(aes(y = ReqAmt),
                               stat = "identity",
                               fill = "tomato") +
                      scale_y_continuous(expand = c(0, 0), 
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(breaks = seq(1, 12, by = 1),
                                           labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                                      "Jul", "Aug", "Sept", "Oct", "Nov", "Dec")) +
                      labs(x = "Declaration Month",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_month

#repeat for obligated funds
obl_fig_month <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentMonth)) +
                      geom_bar(aes(y = OblAmt),
                               stat = "identity",
                               fill = "turquoise4") +
                      scale_y_continuous(expand = c(0, 0),
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(breaks = seq(1, 12, by = 1),
                                           labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                                      "Jul", "Aug", "Sept", "Oct", "Nov", "Dec")) +
                      labs(x = "Declaration Month",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_month

#combine into one plot
plot_col1 <- cowplot::plot_grid(req_fig_month, 
                                     obl_fig_month,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col1

title2 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by declaration month",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_month <- cowplot::plot_grid(
                              title2, plot_col1,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_month

#save file
funds_month_fig_file = here("results","funds-month.png")
ggsave(filename = funds_month_fig_file, plot = funds_fig_month)

#create a table where columns represent requested and obligated funds and rows are 3 most distaster prone-months
decs_month_3 <- EDAdata %>%
                    dplyr::add_count(IncidentMonth) %>%
                    dplyr::filter(dense_rank(-n) < 4) %>%
                    gtsummary::select(IncidentMonth, ReqAmt, OblAmt) %>%
                    gtsummary::tbl_summary(
                      by = IncidentMonth,
                      label = list(IncidentMonth ~ "Month",
                                   ReqAmt ~ "Requested Funds",
                                   OblAmt ~ "Obligated Funds"),
                      statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                    gtsummary::modify_header(label ~ "**Variable**") %>%
                    gtsummary::modify_header(update = list(
                                               stat_1 ~ "**Jan**",
                                               stat_2 ~ "**Sept**",
                                               stat_3 ~ "**Oct**"
                                             )) %>% 
                    gtsummary::modify_spanning_header(c("stat_1", "stat_2", "stat_3") ~ "**Month**") %>%
                    gtsummary::modify_caption("**Table 3. FEMA Funding for the 3 Costliest Months (2012 - 2021)**") %>%
                    gtsummary::bold_labels()
decs_month_3

#save file
decs_month_file = here("results", "decs-month-3.Rds")
saveRDS(decs_month_3, file = decs_month_file)
```

## Predictor: Declaration Type
```{r}
#frequency table since it's categorical
#hide NAs because there are no NAs
summarytools::freq(EDAdata$declarationType, report.nas = FALSE)

#examine graphical relationship with outcomes
#include a jitter function to have a better idea of number of measurements and distribution
#start with obligated amount
obl_fig_DT <- EDAdata %>%
                    ggplot2::ggplot(aes(x = declarationType, y = OblAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "tomato") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      scale_x_discrete(labels = c("Major Disaster Declaration", "Emergency Declaration")) + 
                      labs(x = "Declaration Type",
                           y = "Obligated FEMA Funds",
                           title = "FEMA funds obligated for declaration types")
obl_fig_DT

#save file
obl_DT_fig_file = here("results","obl-DT.png")
ggsave(filename = obl_DT_fig_file, plot = obl_fig_DT)

#repeat for requested funds
req_fig_DT <- EDAdata %>%
                    ggplot2::ggplot(aes(x = declarationType, y = ReqAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "turquoise4") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      scale_x_discrete(labels = c("Major Disaster Declaration", "Emergency Declaration")) + 
                      labs(x = "Declaration Type",
                           y = "Requested FEMA Funds",
                           title = "FEMA funds requested for declaration types")
req_fig_DT

#save file
req_DT_fig_file = here("results","req-DT.png")
ggsave(filename = req_DT_fig_file, plot = req_fig_DT)

#create a table where rows represent requested and obligated funds and columns are declaration type
decs_dtype <- EDAdata %>%
                    gtsummary::select(declarationType, ReqAmt, OblAmt) %>%
                    gtsummary::tbl_summary(
                      by = declarationType,
                      label = list(declarationType ~ "Type",
                                   ReqAmt ~ "Requested Funds",
                                   OblAmt ~ "Obligated Funds"),
                      statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                    gtsummary::modify_header(label ~ "**Variable**") %>% 
                    gtsummary::modify_header(update = list(
                                               stat_1 ~ "**Major Disaster**",
                                               stat_2 ~ "**Emergency**")) %>% 
                    gtsummary::modify_spanning_header(c("stat_1", "stat_2") ~ "**Type**") %>%
                    gtsummary::modify_caption("**Table 4. FEMA Funding for Declaration Types 2012 - 2021**") %>%
                    gtsummary::bold_labels()
decs_dtype

#save file
decs_dtype_file = here("results", "decs-DT.Rds")
saveRDS(decs_dtype, file = decs_dtype_file)
```

## Predictor: Incident Type
```{r}
#frequency table since it's categorical
#hide NAs because there are no NAs
summarytools::freq(EDAdata$incidentType, report.nas = FALSE)

#examine graphical relationship with outcomes
#include a jitter function to have a better idea of number of measurements and distribution
#start with obligated amount
obl_fig_IT <- EDAdata %>%
                    ggplot2::ggplot(aes(x = incidentType, y = OblAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "tomato") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      labs(x = "Incident Type",
                           y = "Obligated FEMA Funds",
                           title = "FEMA funds obligated for incident types") +
                      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
obl_fig_IT

#save file
obl_IT_fig_file = here("results","obl-IT.png")
ggsave(filename = obl_IT_fig_file, plot = obl_fig_IT)

#repeat for requested funds
req_fig_IT <- EDAdata %>%
                    ggplot2::ggplot(aes(x = incidentType, y = ReqAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "turquoise4") +
                      scale_y_continuous(labels = scales::dollar_format()) + 
                      labs(x = "Incident Type",
                           y = "Requested FEMA Funds",
                           title = "FEMA funds requested for declaration types") +
                      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
req_fig_IT

#save file
req_IT_fig_file = here("results","req-IT.png")
ggsave(filename = req_IT_fig_file, plot = req_fig_IT)

#create a table where rows represent requested and obligated funds and columns are declaration type
decs_itype3 <- EDAdata %>%
                    dplyr::add_count(incidentType) %>%
                    dplyr::filter(dense_rank(-n) < 4) %>%
                    gtsummary::select(incidentType, ReqAmt, OblAmt) %>%
                    gtsummary::tbl_summary(
                      by = incidentType,
                      label = list(incidentType ~ "Type",
                                   ReqAmt ~ "Requested Funds",
                                   OblAmt ~ "Obligated Funds"),
                      statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                    gtsummary::modify_header(label ~ "**Variable**") %>% 
                    gtsummary::modify_spanning_header(c("stat_1", "stat_2", "stat_3") ~ "**Type**") %>%
                    gtsummary::modify_caption("**Table 4. FEMA Funding for Declaration Types 2012 - 2021**") %>%
                    gtsummary::bold_labels()
decs_itype3

#save file
decs_itype3_file = here("results", "decs-IT-3.Rds")
saveRDS(decs_itype3, file = decs_itype3_file)
```

## Predictor: Incident Duration
```{r}
#convert class to numeric
EDAdata$IncidentDuration <- as.numeric(EDAdata$IncidentDuration, units="days")

#summary statistics since it's continuous
base::summary(EDAdata$IncidentDuration)
#NAs are ongoing events

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = IncidentDuration)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = IncidentDuration)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$IncidentDuration)

#examine graphical relationship with outcomes of interest
#look at total cost per incident month
req_fig_dur <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentDuration)) +
                      geom_point(aes(y = ReqAmt),
                                 color = "tomato",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Incident Duration (Days)",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_dur

#repeat for obligated funds
obl_fig_dur <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IncidentDuration)) +
                      geom_point(aes(y = OblAmt),
                                 color = "turquoise4",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Incident Duration (Days)",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_dur

#combine into one plot
plot_col2 <- cowplot::plot_grid(req_fig_dur, 
                                     obl_fig_dur,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col2

title3 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by Incident Duration (Under $1B)",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_dur <- cowplot::plot_grid(
                              title3, plot_col2,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_dur

#save file
funds_dur_fig_file = here("results","funds-dur.png")
ggsave(filename = funds_dur_fig_file, plot = funds_fig_dur)
```

## Predictor: Population
```{r}
#convert class to numeric
EDAdata$Population <- as.numeric(gsub(",", "", EDAdata$Population))

#summary statistics since it's continuous
base::summary(EDAdata$Population)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = Population)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = Population)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$Population)

#examine graphical relationship with outcomes of interest
#look at total cost per pop
req_fig_pop <- EDAdata %>%
                    ggplot2::ggplot(aes(x = Population)) +
                      geom_point(aes(y = ReqAmt),
                                 color = "tomato",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(label = comma,
                                         limits = c(0, 40000000)) +
                      labs(x = "State Population",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_pop

#repeat for obligated funds
obl_fig_pop <- EDAdata %>%
                    ggplot2::ggplot(aes(x = Population)) +
                      geom_point(aes(y = OblAmt),
                                 color = "turquoise4",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      scale_x_continuous(label = comma,
                                         limits = c(0, 40000000)) +
                      labs(x = "State Population",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_pop

#combine into one plot
plot_col3 <- cowplot::plot_grid(req_fig_pop, 
                                     obl_fig_pop,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col3

title4 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by State Population (Under $1B)",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_pop <- cowplot::plot_grid(
                              title4, plot_col3,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_pop

#save file
funds_pop_fig_file = here("results","funds-pop.png")
ggsave(filename = funds_pop_fig_file, plot = funds_fig_pop)
```

## Predictor: Total Number of Agencies Involved
```{r}
#summary statistics since it's continuous
base::summary(EDAdata$TotalAgencies)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = TotalAgencies)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = TotalAgencies)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$TotalAgencies)

#examine graphical relationship with outcomes of interest
#look at total cost per incident month
req_fig_ag <- EDAdata %>%
                    ggplot2::ggplot(aes(x = TotalAgencies)) +
                      geom_point(aes(y = ReqAmt),
                                 color = "tomato",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Number of Federal Agencies",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_ag

#repeat for obligated funds
obl_fig_ag <- EDAdata %>%
                    ggplot2::ggplot(aes(x = TotalAgencies)) +
                      geom_point(aes(y = OblAmt),
                                 color = "turquoise4",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Number of Federal Agencies",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_ag

#combine into one plot
plot_col4 <- cowplot::plot_grid(req_fig_ag, 
                                     obl_fig_ag,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col4

title5 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by number of federal agencies (Under $1B)",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_ag <- cowplot::plot_grid(
                              title5, plot_col4,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_ag

#save file
funds_ag_fig_file = here("results","funds-ag.png")
ggsave(filename = funds_ag_fig_file, plot = funds_fig_ag)
```

## Predictor: Number of Counties
```{r}
#summary statistics since it's continuous
base::summary(EDAdata$Counties)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = Counties)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = Counties)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$Counties)

#examine graphical relationship with outcomes of interest
#look at total cost per incident month
req_fig_count <- EDAdata %>%
                    ggplot2::ggplot(aes(x = Counties)) +
                      geom_point(aes(y = ReqAmt),
                                 color = "tomato",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Number of Counties",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_count

#repeat for obligated funds
obl_fig_count <- EDAdata %>%
                    ggplot2::ggplot(aes(x = Counties)) +
                      geom_point(aes(y = OblAmt),
                                 color = "turquoise4",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Number of Counties",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_count

#combine into one plot
plot_col5 <- cowplot::plot_grid(req_fig_count, 
                                     obl_fig_count,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col5

title6 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by number of counties (Under $1B)",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_count <- cowplot::plot_grid(
                              title6, plot_col5,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_count

#save file
funds_count_fig_file = here("results","funds-count.png")
ggsave(filename = funds_count_fig_file, plot = funds_fig_count)
```

## Predictor: FEMA Cost Share Average
```{r}
#summary statistics since it's continuous
base::summary(EDAdata$FEMACSAvg)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = FEMACSAvg)) +
  geom_density()

#normal boxplot
ggplot2::ggplot(data = EDAdata, aes(y = FEMACSAvg)) +
  geom_boxplot()

##QQ plot
car::qqPlot(EDAdata$FEMACSAvg)

#examine graphical relationship with outcomes of interest
#look at total cost per incident month
req_fig_cs <- EDAdata %>%
                    ggplot2::ggplot(aes(x = FEMACSAvg)) +
                      geom_point(aes(y = ReqAmt),
                                 color = "tomato",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Average FEMA Cost Share",
                           y = "Funds ($)",
                           subtitle = "Requested")
req_fig_cs

#repeat for obligated funds
obl_fig_cs <- EDAdata %>%
                    ggplot2::ggplot(aes(x = FEMACSAvg)) +
                      geom_point(aes(y = OblAmt),
                                 color = "turquoise4",
                                 show.legend = FALSE) +
                      scale_y_continuous(expand = c(0, 0),
                                         limits = c(0, 1000000000),
                                         labels = scales::dollar_format()) +
                      labs(x = "Average FEMA Cost Share",
                           y = "Funds ($)",
                           subtitle = "Obligated")
obl_fig_cs

#combine into one plot
plot_col6 <- cowplot::plot_grid(req_fig_cs, 
                                     obl_fig_cs,
                                     ncol = 1,
                                     align = "v",
                                     rel_heights = c(0.5,0.5))
plot_col6

title7 <- cowplot::ggdraw() + 
                    draw_label(
                      "Requested vs obligated FEMA funds by avg FEMA cost share (Under $1B)",
                      fontface = 'bold',
                      x = 0,
                      hjust = 0) +
                    theme( plot.margin = margin(0, 0, 0, 7))

funds_fig_cs <- cowplot::plot_grid(
                              title7, plot_col6,
                              ncol = 1,
                              # rel_heights values control vertical title margins
                              rel_heights = c(0.1, 1))
funds_fig_cs

#save file
funds_cs_fig_file = here("results","funds-cs.png")
ggsave(filename = funds_cs_fig_file, plot = funds_fig_cs)
```

## Predictor: IH Program
```{r}
#convert to factor
EDAdata$IH <- as.factor(EDAdata$IH)

#frequency table since it's categorical
#hide NAs because there are no NAs
summarytools::freq(EDAdata$IH, report.nas = FALSE)

#examine graphical relationship with outcomes
#include a jitter function to have a better idea of number of measurements and distribution
#start with obligated amount
obl_fig_IH <- EDAdata %>%
                    ggplot2::ggplot(aes(x = IH, y = OblAmt)) +
                      geom_boxplot() +
                      geom_jitter(alpha = 0.5, width = 0.2, height = 0.2, color = "tomato") +
                      scale_y_continuous(labels = scales::dollar_format()) +
                      scale_x_discrete(labels = c("Not Awarded", "Awarded")) +
                      labs(x = "IH Program",
                           y = "Obligated FEMA Funds",
                           title = "FEMA funds obligated when IH program is awarded") +
                      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
obl_fig_IH
